The Magnetorotational Instability Prefers Three Dimensions

### 2019-06-24 paper/paper.tex
Updated figures; added to text a bit.

@@

We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z) + \gamma t}$. This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework. The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions on all ten variables.

@@

\begin{figure}[h]
  \includegraphics[width=0.33\textwidth]{run_11_output_growthrates.png}
  \includegraphics[width=0.33\textwidth]{run_12_output_growthrates.png}
  \includegraphics[width=0.33\textwidth]{run_13_output_growthrates.png}
  \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
  \label{fig:growth_rate}
\end{figure}

Numerical simulations have consistently axisymmetric modes dominating early evolution of the MRI (cite Stone 95, etc) before parasitic modes cause a breakdown into MHD turbulence. However, our results recover this limit, as $\SSC$ becomes large. In this case

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{alpha_vs_ssc.png}
  \caption{Angle $\alpha$ vs $\SSC$.}
  \label{fig:alpha}
\end{figure}
\section{Non-Normality}
\label{sec:non-normality}

The linear operator for the MRI, like other shearing systems, is non-normal. 
%If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}

>>>>

\begin{itemize}
\item 3D $\to$ dynamo
\item stars, NSSL
\item experiments
\end{itemize}

@@

We solve equations~(\ref{eq:ns})--~(\ref{eq:divu}) by assuming harmonic perturbations in $y$ and $z$, $f = \hat{f}(x) e^{i(k_y y + k_z z) + \gamma t}$. This reduces to an eigenvalue problem in $x$, which we solve using the \emph{Dedalus} pseudospectral framework. The advantage of this particular form is that it can be expressed as a set of ten first-order equations with Dirichlet boundary conditions for either no-slip or stress-free boundary conditions.

Our interest is in nearly ideal ($\eta = 0$), inviscid ($\nu = 0$) conditions; we thus set $\eta=\nu=10^{-5}$, though we have confirmed our main results are insensitive to this particular value.

@@

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{run_11_output_growthrates.png}
  \includegraphics[width=\columnwidth]{run_12_output_growthrates.png}
  \includegraphics[width=\columnwidth]{run_13_output_growthrates.png}
  \caption{Growth rate as a function of $k_z$ and $k_y$. The maximum growth rate is far from $k_y = 0$, indicating the MRI is most unstable in a fully three-dimensional configuration.}
  \label{fig:growth_rate}
\end{figure}

Numerical simulations have consistently axisymmetric modes dominating early evolution of the MRI (cite Stone 95, etc) before parasitic modes cause a breakdown into MHD turbulence. However, our results recover this limit, as $\SSC$ becomes large. Figure~\ref{fig:alpha} plots $\alpha = \arctan k_y/k_z$ as  function of $\SSC$. Above $\SSC \gtrsim 2$, $\alpha$ is zero, indicating that axisymmetric modes have the fastest growth rates.

\begin{figure}[h]
  \includegraphics[width=\columnwidth]{alpha_vs_ssc_grid.pdf}
  \caption{Angle $\alpha$ vs $\SSC$. Stability limit is at $\SSC \simeq 0.102$}
  \label{fig:alpha}
\end{figure}
\section{Non-Normality}
\label{sec:non-normality}

\section{Speculations}
\label{sec:speculations}

\begin{itemize}
\item non-normal and $k_y$ even when $\alpha = 0$ implies important role for linear dynamics at early times. 
\item Following Squire + Bhattacharjee, this suggests that important stellar layers may in fact be driven by linear dynamics
\item why is MRI non-oscillatory when hydro shear is oscil or stable?
\end{itemize}
The linear operator for the MRI, like other shearing systems, is non-normal. 
%If in two-column mode, this environment will change to single-column
% format so that long equations can be displayed. Use
% sparingly.
%\begin{widetext}
% put long equation here
%\end{widetext}
